---
title: "CTC Analysis"
author: "Nikhil Kalra"
format: 
  docx:
    #reference-doc: "PhD-thesis_NK_2024-11"
    toc: true
  pdf:
    toc: true
editor: visual
execute:
  echo: true
  warning: false
  output: asis
fig-dpi: 300
fig-width: 6
fig-height: 8
tbl-cap-location: top
---

# Loading packages

```{r}
pacman::p_load(conflicted, flextable, formattable, ggbeeswarm,
               ggalluvial, ggbreak, here, readxl, scales, tidyverse,
               wrappedtools)
conflicts_prefer(dplyr::filter)
conflicts_prefer(formattable::comma)
conflicts_prefer(readxl::read_xlsx)

```

# Importing data

```{r}
biopsydata <- read_xlsx("Data/CTCanalysis_NK.xlsx")
baselinedata <- read_xlsx("Data/CTCanalysis_NK.xlsx", sheet = 2)
qualvars <- ColSeeker(biopsydata,
                     namepattern=c("^C","meno","Stanze","Ne"))
```

# Data Cleaning

## Mutating Pathology column

```{r}
rawdata <- 
  biopsydata |> 
  mutate(
    patho_ER = case_when(
      str_detect(Pathology, ".*?ER\\s*[><=]?\\s*\\d+%?.*") ~
        str_replace(Pathology, ".*?(ER\\s*[><=]?\\s*\\d+%?).*", "\\1") |> 
        str_replace("ER\\s*[><=]?\\s*", "")
    ),
    patho_PR = case_when(
      str_detect(Pathology, ".*?PR\\s*[><=]?\\s*\\d+%?.*") ~
        str_replace(Pathology, ".*?(PR\\s*[><=]?\\s*\\d+%?).*", "\\1") |> 
        str_replace("PR\\s*[><=]?\\s*", "")
    ),
    patho_HER2 = case_when(
      str_detect(Pathology, ".*HER2 \\D{3}.*") ~
        str_replace(Pathology, ".*(HER2 \\D{3}).*", "\\1") |> 
        str_replace("HER2 ", "")
    ),
    patho_Ki67 = case_when(
      str_detect(Pathology, "Ki67 <?\\d+%") ~
        str_replace(Pathology, ".*(Ki67 <?\\d+%).*", "\\1") |> 
        str_replace_all("Ki67\\s*|[<%]", "")
    )
  ) |> 
  mutate(
    across(all_of(qualvars$names), factor) # Convert specific columns to factors
  )

```

## Mutating Staging column

```{r}
identifiers <- c("T", "N", "M", "G", "L", "V", "n", "R")

# Extract columns dynamically and automate column naming
rawdata <- rawdata %>%
  mutate(Staging = replace_na(Staging, "")) %>%  # Replace NA with empty strings
  bind_cols(
    map_dfc(identifiers, ~ tibble(!!paste0(.x, "_stage") := str_extract(rawdata$Staging, paste0(.x, "[A-Za-z0-9]{1}"))))
  ) %>%
  mutate(across(ends_with("_stage"), ~ str_replace_all(., "^.*?(\\w)$", "\\1")))  # Extract only single character

```

# Selecting cases for analysis

1.  Cases where Epithelial cells were detected have to excluded from the study
2.  Cases with incomplete/damaged/discarded blood have to excluded as well

```{r}
rawdata <- 
  rawdata |> 
  mutate(
    complete=case_when(
      # Stanze_type != "Not_received" &
      #   Cell_Category != "Epithelial"~ "yes",
      Stanze_type == "Not_received" | 
        Cell_Category == "Epithelial" ~ "no",
      .default = "yes"),
    CApositive=case_when(
      complete == "yes" &
        Cancer == "positive" ~ "yes",
      complete == "yes" &
        Cancer == "negative" ~ "no"))

#Summary_table for samples included in analysis
summary_data <- rawdata |>
  group_by(Cancer) |>
  summarise(
    overall_cases = n(),  # Total cases
    selected_cases = sum(complete == "yes", na.rm = TRUE)) |>  # Selected cases
  mutate(
    overall_freq = formattable::percent(overall_cases / sum(overall_cases)),
    selected_freq = formattable::percent(selected_cases / sum(selected_cases)))

bind_rows(
  summary_data,
  tibble(
    Cancer = "Total",
    overall_cases = sum(summary_data$overall_cases, na.rm = TRUE),
    selected_cases = sum(summary_data$selected_cases, na.rm = TRUE),
    overall_freq = formattable::percent(1),  # Total frequency = 100%
    selected_freq = formattable::percent(1)  # Total selected frequency = 100%
  )
) |> flextable()

```

# Table 1 : Clinicopathological parameters

```{r}
# Age plot
rawdata |>
  filter(complete=="yes") |>
  select(Age, Cancer) |> 
  filter(!is.na(Age)) |> 
  ggplot(aes(Age, color=Cancer)) + geom_density()

# p-value: Age
## Normality test
rawdata |> 
  filter(complete == "yes") |>
  group_by(Cancer) |> 
  summarise(cases = n(),
            shapiro_pvalue = shapiro.test(Age)$p.value)

## Wilcoxon test
rawdata |> 
  filter(complete == "yes") |>
  wilcox.test(Age ~ Cancer, data = _)

# CTC positive cases before biopsy
rawdata |> 
  filter(complete == "yes") |> 
  group_by(Cancer) |> 
  summarise(
    Method01 = list(Method01_pre[Method01_pre > 0]),
    Method02 = list(Method02_pre[Method02_pre > 0])
  ) |> 
 summarise(
    shapiro_CTCs_Method01 = shapiro.test(unlist(Method01))$p.value,
    shapiro_CTCs_Method02 = shapiro.test(unlist(Method02))$p.value,
    wilcox_pvalue = wilcox.test(unlist(Method01), unlist(Method02), exact = FALSE)$p.value
  )

# CTC counts before biopsy
rawdata |> 
  filter(complete == "yes") |> 
  select(Patient_ID, Cancer, Method01_pre, Method02_pre) |> 
  group_by(Cancer) |> 
  summarise(
    across(
      contains("Method"),
      ~ sum(.x, na.rm = TRUE),
      .names = "{.col}_sum"
    ))

# Clinical variables
# Initialize a list to store results
table1 <- list()

# Total counts
table1$total <- nrow(rawdata |> filter(complete=="yes"))

# Breast cancer status counts
table1$cancer <- rawdata |> 
  filter(complete=="yes") |> 
  group_by(Cancer)  |> 
  summarise(cases = n()) |> 
  mutate(freq = formattable::percent(cases / sum(cases)))

# Age at diagnosis: median, min, max
table1$age <- rawdata |> 
  filter(complete=="yes") |> 
  group_by(Cancer) |>
  summarise(
    median = median(Age, na.rm = TRUE),
    min = min(Age, na.rm = TRUE),
    max = max(Age, na.rm = TRUE))

# Menopause status
table1$menopause <- rawdata |> 
  filter(complete == "yes") |> 
  mutate(
    Menopause = ifelse(Menopause == "perimenopausal", "postmenopausal", Menopause)
  ) |> 
  # Calculate global totals for CA groups
  mutate(
    total_CAnegative = sum(Cancer == "negative", na.rm = TRUE),
    total_CApositive = sum(Cancer == "positive", na.rm = TRUE)
  ) |> 
  group_by(Menopause) |> 
  summarise(
    negative = sum(Cancer == "negative", na.rm = TRUE),
    positive = sum(Cancer == "positive", na.rm = TRUE),
    total_CAnegative = first(total_CAnegative),  # Use global totals for CA-negative
    total_CApositive = first(total_CApositive),  # Use global totals for CA-positive
    CA_negative = paste0(negative, " (", formattable::percent(negative / total_CAnegative), ")"),
    CA_positive = paste0(positive, " (", formattable::percent(positive / total_CApositive), ")")
  ) |> 
  select(Menopause, CA_positive, CA_negative)

# Tissue biopsy
table1$biopsy <- rawdata |> 
  filter(complete=="yes") |> 
  mutate(
    total_negative = sum(Cancer == "negative", na.rm = TRUE),
    total_positive = sum(Cancer == "positive", na.rm = TRUE)
  ) |> 
  group_by(Stanze_type) |> 
  summarise(
    negative = sum(Cancer == "negative", na.rm = TRUE),
    positive = sum(Cancer == "positive", na.rm = TRUE),
    total_CAnegative = first(total_negative),
    total_CApositive = first(total_positive),
    CA_negative = paste0(negative," (", formattable::percent(negative/total_CAnegative),")"),
    CA_positive = paste0(positive," (", formattable::percent(positive/total_CApositive),")")) |> 
  select(Stanze_type, CA_positive, CA_negative)

# Tissue histology
rawdata |> 
  filter(complete == "yes", Cancer=="positive") |> 
  select(Cancer, Histology) |> 
  mutate(
    CApos_hist = case_when(
      grepl("IDC|NST", Histology, ignore.case = TRUE) ~ "IDC/NST",
      grepl("DCIS", Histology, ignore.case = TRUE) ~ "DCIS",
      is.na(Histology) ~ "NA",
      TRUE ~ Histology)) |> 
  group_by(CApos_hist) |> 
  summarise(
    Cases = n()) |> 
  ungroup() |> 
  mutate(Percentage = (Cases / sum(Cases)) * 100)

rawdata |> 
  filter(complete=="yes", Cancer=="negative") |> 
  select(Cancer, Histology) |>  
  mutate(
    CAneg_hist = case_when(
      grepl("Fibrosis|Periductal Fibrosis|Fibrous Mastopathy|Cyst|Benign Cyst| Fibromatosis|Necrotic tissue", Histology, ignore.case = TRUE) ~ "Fibrous Lesions",
      grepl("Fibroadenoma|Fibroma|Hemartom", Histology, ignore.case = TRUE) ~ "Fibrous Tumors",
      grepl("Intraductal Papilloma|UDH|CLIS", Histology, ignore.case = TRUE) ~ "Epithelial lesions",
      is.na(Histology) ~ "NA",
      TRUE ~ Histology)) |> 
  group_by(CAneg_hist) |> 
  summarise(
    Cases = n()) |> 
  ungroup() |> 
  mutate(Percentage = (Cases / sum(Cases)) * 100)

# Calculating Table 1 values for ER, PR, HER2, Ki67%
# ER_status
table1$ER_status <- rawdata |> 
  filter(complete=="yes") |> 
  mutate(
    ER_status = case_when(
      patho_ER == "0" ~ "0",
      is.na(patho_ER) ~ "NA",
      as.numeric(patho_ER) > 0 ~ ">0"
      # .default = ">0"
      # TRUE ~ ">0"
      )) |> 
  group_by(ER_status) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(ER_status, CA_positive, CA_negative)

# PR_status
table1$PR_status <- rawdata |> 
  filter(complete=="yes") |> 
  mutate(
    PR_status = case_when(
      patho_PR == "0" ~ "0",
      is.na(patho_PR) ~ "NA",
      as.numeric(patho_PR) > 0 ~ ">0"
      # .default = ">0"
      # TRUE ~ ">0"
      )) |> 
  group_by(PR_status) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(PR_status, CA_positive, CA_negative)

# HER2_status
table1$HER2_status <- rawdata |> 
    filter(complete=="yes") |> 
    group_by(patho_HER2) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(patho_HER2, CA_positive, CA_negative)

# Ki67_status
table1$Ki67_status <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
    patho_Ki67 = case_when(
      patho_Ki67 <= 15 ~ "<=15% (low)",
      patho_Ki67 > 15 ~ ">15% (high)",
      is.na(patho_Ki67) ~ "NA"
      # .default = ">0"
      # TRUE ~ ">0"
      )) |>
    group_by(patho_Ki67) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(patho_Ki67, CA_positive, CA_negative)

# Molecular subtype
rawdata |> 
  # Convert numeric columns while leaving patho_HER2 as character
  mutate(across(c(patho_ER, patho_PR, patho_Ki67), ~ as.numeric(.))) |> 
  filter(complete == "yes", Cancer=="negative") |> 
  select(Cancer, contains("patho_")) |>  
  # Apply the subtyping logic
  mutate(
    Subtype = case_when(
      # Luminal A
      (patho_ER > 0 | patho_PR > 0) & patho_HER2 == "neg" & patho_Ki67 <= 15 ~ "Luminal_A",
      # Luminal B
      (patho_ER > 0 | patho_PR > 0) & (patho_HER2 %in% c("neg", "pos")) & patho_Ki67 > 15 ~ "Luminal_B",
      # HER2-enriched
      patho_ER == 0 & patho_PR == 0 & patho_HER2 == "pos" & patho_Ki67 >= 0 ~ "HER2/neu",
      # Triple Negative Breast Cancer (TNBC)
      patho_ER == 0 & patho_PR == 0 & patho_HER2 == "neg" & patho_Ki67 >= 0 ~ "TNBC",
      # patho_ER == 0 & patho_PR == 0 & patho_HER2 == "neg" & (patho_Ki67 > 15 & patho_Ki67 <= 30) ~ "TNBC",
      # # Basal-like (Subtype of TNBC)
      # patho_ER == 0 & patho_PR == 0 & patho_HER2 == "neg" & patho_Ki67 > 30 ~ "Basal-like",
      # # Claudin-low (Subtype of TNBC, stem-like features)
      # patho_ER == 0 & patho_PR == 0 & patho_HER2 == "neg" & (patho_Ki67 >= 0 & patho_Ki67 <= 15) ~ "Claudin-low",
      # Triple Positive
      (patho_ER > 0 | patho_PR > 0) & patho_HER2 == "pos" & patho_Ki67 >= 0 ~ "Triple_Positive",
      # Normal-like
      # (patho_ER > 0 | patho_PR > 0) & patho_HER2 == "neg" & patho_Ki67 <= 15 ~ "Normal-like",
      # Default for unmatched cases
      TRUE ~ "NA")) |> # view()
  group_by(Subtype, Cancer) |> 
  summarise(Cases = n()) |> 
  ungroup() |> 
  mutate(Percentage = (Cases / sum(Cases)) * 100)

# Staging: T, N, Metastasis
table1$T_stage <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
    T_stage = case_when(
      T_stage == "i"  ~ "Tis",
      (T_stage == 1 | T_stage == 2) ~ "T1+T2",
      (T_stage == 3 | T_stage == 4) ~ "T3+T4",
      is.na(T_stage) ~ "NA"
      # .default = ">0"
      # TRUE ~ ">0"
      )) |>
    group_by(T_stage) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(T_stage, CA_positive, CA_negative)

# N stage
table1$N_stage <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
    N_stage = case_when(
       N_stage == 0 ~ "N0", 
       N_stage == "x" ~ "Nx",
       N_stage == 1 ~ "N1",
      is.na(N_stage) ~ "NA")) |>
    group_by(N_stage) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(N_stage, CA_positive, CA_negative)

# M stage
table1$M_stage <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
    M_stage = case_when(
      M_stage == 0 ~ "M0",
      M_stage == "x" | M_stage == "X" ~ "Mx",
      M_stage == 1 ~ "M1",
      is.na(M_stage) ~ "NA" )) |>
    group_by(M_stage) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(M_stage, CA_positive, CA_negative)

# Tumor grade
table1$G_stage <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
    G_stage = case_when(
      (G_stage == 1 | G_stage == 2) ~ "G1+G2",
      (G_stage == 3) ~ "G3",
      is.na(G_stage) ~ "NA"
      )) |>
    group_by(G_stage) |> 
    summarise(negative = sum(Cancer == "negative", na.rm = TRUE),
              positive = sum(Cancer == "positive", na.rm = TRUE)) |>
    mutate(total_negative = sum(negative),
           total_positive = sum(positive),
           CA_negative = paste0(negative, " (", formattable::percent(negative / total_negative), ")"),
           CA_positive = paste0(positive, " (", formattable::percent(positive / total_positive), ")")) |>
    select(G_stage, CA_positive, CA_negative)

# Global function for all clinical parameters
variables <- c("L_stage", "V_stage", "n_stage", "R_stage")

# if (!exists("table1")) table1 <- list()
# Loop through variables and assign results to table1
for (var in variables) {
  table1[[var]] <- rawdata |> 
    filter(complete=="yes") |> 
    mutate(
      Group = case_when(
        .data[[var]] == "0" ~ "0",
        is.na(.data[[var]]) ~ "NA",
        TRUE ~ "1"
      )
    ) |>
    group_by(Group) |>
    summarise(
      negative = sum(Cancer == "negative", na.rm = TRUE),
      positive = sum(Cancer == "positive", na.rm = TRUE)
    ) |>
    mutate(
      total_negative = sum(negative),
      total_positive = sum(positive),
      CA_negative = paste0(negative, " (",
                           formattable::percent(negative / total_negative), ")"),
      CA_positive = paste0(positive, " (",
                           formattable::percent(positive / total_positive), ")")
    ) |>
    select(Group, CA_positive, CA_negative)
}

print(names(table1))

```

# Table 2: CTC counts from both methods
## Overall cases:

```{r}
## Before biopsy
summary_data <- rawdata |>  
  filter(complete == "yes") |> 
  select(Patient_ID, Cancer, contains("pre"), -Method01_preII) |> 
  summarise(across(
    starts_with("Method"),
    list(
      sum = ~sum(.x, na.rm = TRUE),
      positive_cases = ~sum(.x > 0, na.rm = TRUE),
      median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
      range = ~if (any(.x > 0, na.rm = TRUE)) {
        paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")
      } else {
        NA_character_
      }
    ),
    .names = "{.col}_{.fn}"  # Dynamically name columns
  ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("CTCs_pre_biopsy"),
    values_to = "Value"
  ) |> 
# pivot_wider( names_from = CTCs_pre_biopsy,
#             values_from = Summary) 
  flextable() |> 
  set_caption(caption = "Overall cases: CTCs Pre-Biopsy") |> 
  set_table_properties(width = 0.8, layout = "autofit")


## After biopsy
summary_data <- rawdata |>  
      filter(complete=="yes") |> 
      select(Patient_ID, Cancer, contains("post")) |>
      summarise(across(
        starts_with("Method"),
        list(
      sum = ~sum(.x, na.rm = TRUE),
      positive_cases = ~sum(.x > 0, na.rm = TRUE),
      median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
      range = ~if (any(.x > 0, na.rm = TRUE)) {
        paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")  
        # Range for positive cases only
      } else {
        NA_character_
      }
    ),
    .names = "{.col}_{.fn}"  # Dynamically name columns
  ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("CTCs_post_biopsy"),
    values_to = "Value"
  ) |> 
# pivot_wider( names_from = CTCs_pre_biopsy,
#             values_from = Summary) 
  flextable() |> 
  set_caption(caption = "Overall cases: CTCs Post-Biopsy") |> 
  set_table_properties(width = 0.8, layout = "autofit")


## Combined: Before + After biopsy
summary_data <- rawdata |>  
      filter(complete=="yes") |> 
      mutate(Method01_combined = Method01_pre + Method01_post,
             Method02_combined = Method02_pre + Method02_post) |> 
      select(Patient_ID, Cancer, contains("Method"), -Method01_preII) |>
      summarise(across(
        contains("combined"),
        list(
              sum = ~sum(.x, na.rm = TRUE),
              positive_cases = ~sum(.x > 0, na.rm = TRUE),
              median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
              range = ~if (any(.x > 0, na.rm = TRUE)) {
                paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")  
                # Range for positive cases only
              } else {
                NA_character_
              }
            ),
            .names = "{.col}_{.fn}"  # Dynamically name columns
          ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("Combined_CTCs"),
    values_to = "Value"
  ) |>
  flextable() |> 
  set_caption(caption = "Overall cases: Combined CTCs (pre- & post-biopsy)") |> 
  set_table_properties(width = 0.8, layout = "autofit")

```

## Overall cases: Wilcoxon, McNemar, Kendall's tau, Cohen's kappa

```{r}
# Load necessary libraries
# install.packages("irr")
# install.packages("vcd")
# install.packages("tau")

# Load necessary libraries
library(stats) # For Wilcoxon test
library(irr)   # For Cohen's kappa
library(vcd)   # For McNemar test
library(tau)   # For Kendall's tau

# Example data from the table
Method01_ctc <- list(
  pre = rawdata |> filter(complete=="yes" ) |> pull(Method01_pre),
  post = rawdata |> filter(complete=="yes") |> pull(Method01_post),
  combined = rawdata |> filter(complete=="yes") |> mutate(Method01_combined = Method01_pre + Method01_post) |> pull(Method01_combined))

Method02_ctc <- list(
  pre = rawdata |> filter(complete=="yes") |> pull(Method02_pre),
  post = rawdata |> filter(complete=="yes") |> pull(Method02_post),
  combined = rawdata |> filter(complete=="yes") |> mutate(Method02_combined = Method02_pre + Method02_post) |> pull(Method02_combined))

## Binary outcomes (positive or negative outcome) for each method
Metho01_binary <- list(
  pre = ifelse(Method01_ctc$pre > 0, 1, 0),
  post = ifelse(Method01_ctc$post > 0, 1, 0),
  combined = ifelse(Method01_ctc$combined > 0, 1, 0))

Method02_binary <- list(
  pre = ifelse(Method02_ctc$pre > 0, 1, 0),
  post = ifelse(Method02_ctc$post > 0, 1, 0),
  combined = ifelse(Method02_ctc$combined > 0, 1, 0))

# Ensure both datasets are paired (same number of observations)
if (length(Metho01_binary$pre) != length(Method02_binary$pre)) {
  stop("The data lengths for 'pre' measurements do not match!")
}

if (length(Metho01_binary$post) != length(Method02_binary$post)) {
  stop("The data lengths for 'post' measurements do not match!")
}

if (length(Metho01_binary$combined) != length(Method02_binary$combined)) {
  stop("The data lengths for 'post' measurements do not match!")
}

# Wicoxon Test
wilcoxon_pre <- wilcox.test(Method01_ctc$pre, Method02_ctc$pre, paired = TRUE, exact=FALSE)$p.value
wilcoxon_post <- wilcox.test(Method01_ctc$post, Method02_ctc$post, paired = TRUE, exact=FALSE)$p.value
wilcoxon_combined <- wilcox.test(Method01_ctc$combined, Method02_ctc$combined, paired = TRUE, exact=FALSE)$p.value

# McNemar Test
# prepare tables for McNemar's test
table_pre <- table(Metho01_binary$pre, Method02_binary$pre)
table_post <- table(Metho01_binary$post, Method02_binary$post)
table_combined <- table(Metho01_binary$combined, Method02_binary$combined)

# McNemar p-values
McNemar_pre <- mcnemar.test(table_pre)$p.value
McNemar_post <- mcnemar.test(table_post)$p.value
McNemar_combined <- mcnemar.test(table_combined)$p.value

# Kendall's Tau estimates: Correlation between CTC numbers
tau_pre <- cor.test(Method01_ctc$pre, Method02_ctc$pre, method="kendall")
tau_post <- cor.test(Method01_ctc$post, Method02_ctc$post, method="kendall")
tau_combined <- cor.test(Method01_ctc$combined, Method02_ctc$combined, method="kendall")

# Cohen's kappa for Pre- and Post-biopsy
kappa_pre <- kappa2(cbind(Metho01_binary$pre, Method02_binary$pre))
kappa_post <- kappa2(cbind(Metho01_binary$post, Method02_binary$post))
kappa_combined <- kappa2(cbind(Metho01_binary$combined, Method02_binary$combined))

# print results
data.frame(
  Statistic = c("Wilcoxon", "McNemar", "Kendall's Tau", "Cohen's Kappa"),
  Pre = c(
    round(wilcoxon_pre, 4),
    round(McNemar_pre, 4),
    paste0(round(tau_pre$estimate, 4), " (p=", round(tau_pre$p.value, 4), ")"),
    paste0(round(kappa_pre$value, 4), " (p=", round(kappa_pre$p.value, 4), ")")),
  Post = c(
    round(wilcoxon_post, 4),
    round(McNemar_post, 4),
    paste0(round(tau_post$estimate, 4), " (p=", round(tau_post$p.value, 4), ")"),
    paste0(round(kappa_post$value, 4), " (p=", round(kappa_post$p.value, 4), ")")),
  Combined = c(
    round(wilcoxon_combined, 4),
    round(McNemar_combined, 4),
    paste0(round(tau_combined$estimate, 4), " (p=", round(tau_combined$p.value, 4), ")"),
    paste0(round(kappa_combined$value, 4), " (p=", round(kappa_combined$p.value, 4), ")"))) |>
  flextable() |>
    set_caption(caption = "Overall cases: Pre-, Post- and Combined Analysis") |>
    autofit()


## Method02: Prepare contingency matrices for pre and post
# contingency_pre <- matrix(c(sum(Metho01_binary$pre == 1 & Method02_binary$pre == 1),
#                          sum(Metho01_binary$pre == 1 & Method02_binary$pre == 0),
#                          sum(Metho01_binary$pre == 0 & Method02_binary$pre == 1),
#                          sum(Metho01_binary$pre == 0 & Method02_binary$pre == 0)),
#                          nrow = 2)
# 
# contingency_post <- matrix(c(sum(Metho01_binary$post == 1 & Method02_binary$post == 1),
#                          sum(Metho01_binary$post == 1 & Method02_binary$post == 0),
#                          sum(Metho01_binary$post == 0 & Method02_binary$post == 1),
#                          sum(Metho01_binary$post == 0 & Method02_binary$post == 0)),
#                          nrow = 2)
# contingency_combined <- matrix(c(sum(Metho01_binary$combined == 1 &                         #                                      Method02_binary$combined == 1),
#                          sum(Metho01_binary$combined == 1 & Method02_binary$combined == 0),
#                          sum(Metho01_binary$combined == 0 & Method02_binary$combined == 1),
#                          sum(Metho01_binary$combined == 0 & Method02_binary$combined == 0)),
#                          nrow = 2)
```

## Malignant cases:

```{r}
## Before biopsy
summary_data <- rawdata |>
  filter(complete == "yes" & CApositive == "yes") |>
  select(Patient_ID, Cancer, contains("pre"), -Method01_preII) |> 
  summarise(across(
    starts_with("Method"),
    list(
      sum = ~sum(.x, na.rm = TRUE),
      positive_cases = ~sum(.x > 0, na.rm = TRUE),
      median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
      range = ~if (any(.x > 0, na.rm = TRUE)) {
        paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")
      } else {
        NA_character_
      }
    ),
    .names = "{.col}_{.fn}"  # Dynamically name columns
  ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("CTCs_pre_biopsy"),
    values_to = "Value"
  ) |> 
# pivot_wider( names_from = CTCs_pre_biopsy,
#             values_from = Summary) 
  flextable() |> 
  set_caption(caption = "Breast cancer positive: CTCs pre-Biopsy") |> 
  set_table_properties(width = 0.8, layout = "autofit")


## After biopsy
summary_data <- rawdata |>  
      filter(complete == "yes" & CApositive == "yes") |>
      select(Patient_ID, Cancer, contains("post")) |>
      summarise(across(
        starts_with("Method"),
        list(
      sum = ~sum(.x, na.rm = TRUE),
      positive_cases = ~sum(.x > 0, na.rm = TRUE),
      median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
      range = ~if (any(.x > 0, na.rm = TRUE)) {
        paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")  
        # Range for positive cases only
      } else {
        NA_character_
      }
    ),
    .names = "{.col}_{.fn}"  # Dynamically name columns
  ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("CTCs_post_biopsy"),
    values_to = "Value"
  ) |> 
# pivot_wider( names_from = CTCs_pre_biopsy,
#             values_from = Summary) 
  flextable() |> 
  set_caption(caption = "Breast cancer positive: CTCs post-Biopsy") |> 
  set_table_properties(width = 0.8, layout = "autofit")

## Combined: Before + After biopsy
summary_data <- rawdata |>  
      filter(complete == "yes" & CApositive == "yes") |>
      mutate(Method01_combined = Method01_pre + Method01_post,
             Method02_combined = Method02_pre + Method02_post) |> 
      select(Patient_ID, Cancer, contains("Method"), -Method01_preII) |>
      summarise(across(
        contains("combined"),
        list(
              sum = ~sum(.x, na.rm = TRUE),
              positive_cases = ~sum(.x > 0, na.rm = TRUE),
              median = ~median(.x[.x > 0], na.rm = TRUE),  # Median for positive cases only
              range = ~if (any(.x > 0, na.rm = TRUE)) {
                paste0(range(.x[.x > 0], na.rm = TRUE), collapse = " - ")  
                # Range for positive cases only
              } else {
                NA_character_
              }
            ),
            .names = "{.col}_{.fn}"  # Dynamically name columns
          ))

# Reshape the result to a long table with rows for Method01 and Method02
summary_data %>%
  mutate(across(everything(), as.character)) |> 
  pivot_longer(
    cols = everything(),
    names_to = c("Combined_CTCs"),
    values_to = "Value"
  ) |>
  flextable() |> 
  set_caption(caption = "Breast cancer positive: Combined CTCs (pre- & post-biopsy)") |> 
  set_table_properties(width = 0.8, layout = "autofit")

```

## Malignant cases: Wilcoxon, McNemar, Kendall's tau, Cohen's kappa

```{r}
Method01_ctc <- list(
  pre = rawdata |> filter(complete=="yes" & CApositive == "yes") |> pull(Method01_pre),
  post = rawdata |> filter(complete=="yes" & CApositive == "yes") |> pull(Method01_post),
  combined = rawdata |> 
             filter(complete=="yes" & CApositive == "yes") |>                                          mutate(Method01_combined = Method01_pre + Method01_post) |>                               pull(Method01_combined))

Method02_ctc <- list(
  pre = rawdata |> filter(complete=="yes" & CApositive == "yes") |> pull(Method02_pre),
  post = rawdata |> filter(complete=="yes" & CApositive == "yes") |> pull(Method02_post),
  combined = rawdata |> 
             filter(complete=="yes" & CApositive == "yes") |> 
             mutate(Method02_combined = Method02_pre + Method02_post) |>                               pull(Method02_combined))

## Binary outcomes (positive or negative outcome) for each method
Metho01_binary <- list(
  pre = ifelse(Method01_ctc$pre > 0, 1, 0),
  post = ifelse(Method01_ctc$post > 0, 1, 0),
  combined = ifelse(Method01_ctc$combined > 0, 1, 0))

Method02_binary <- list(
  pre = ifelse(Method02_ctc$pre > 0, 1, 0),
  post = ifelse(Method02_ctc$post > 0, 1, 0),
  combined = ifelse(Method02_ctc$combined > 0, 1, 0))

# Ensure both datasets are paired (same number of observations)
if (length(Metho01_binary$pre) != length(Method02_binary$pre)) {
  stop("The data lengths for 'pre' measurements do not match!")
}

if (length(Metho01_binary$post) != length(Method02_binary$post)) {
  stop("The data lengths for 'post' measurements do not match!")
}

if (length(Metho01_binary$combined) != length(Method02_binary$combined)) {
  stop("The data lengths for 'post' measurements do not match!")
}

wilcoxon_pre <- wilcox.test(Method01_ctc$pre, Method02_ctc$pre, paired = TRUE, exact=FALSE)$p.value
wilcoxon_post <- wilcox.test(Method01_ctc$post, Method02_ctc$post, paired = TRUE, exact=FALSE)$p.value
wilcoxon_combined <- wilcox.test(Method01_ctc$combined, Method02_ctc$combined, paired = TRUE, exact=FALSE)$p.value

# McNemar Test
# prepare tables for McNemar's test
table_pre <- table(Metho01_binary$pre, Method02_binary$pre)
table_post <- table(Metho01_binary$post, Method02_binary$post)
table_combined <- table(Metho01_binary$combined, Method02_binary$combined)

# McNemar p-values
McNemar_pre <- mcnemar.test(table_pre)$p.value
McNemar_post <- mcnemar.test(table_post)$p.value
McNemar_combined <- mcnemar.test(table_combined)$p.value

# Kendall's Tau estimates: Correlation between CTC numbers
tau_pre <- cor.test(Method01_ctc$pre, Method02_ctc$pre, method="kendall")
tau_post <- cor.test(Method01_ctc$post, Method02_ctc$post, method="kendall")
tau_combined <- cor.test(Method01_ctc$combined, Method02_ctc$combined, method="kendall")

# Cohen's kappa for Pre- and Post-biopsy
kappa_pre <- kappa2(cbind(Metho01_binary$pre, Method02_binary$pre))
kappa_post <- kappa2(cbind(Metho01_binary$post, Method02_binary$post))
kappa_combined <- kappa2(cbind(Metho01_binary$combined, Method02_binary$combined))

# print results
data.frame(
  Statistic = c("Wilcoxon", "McNemar", "Kendall's Tau", "Cohen's Kappa"),
  Pre = c(
    round(wilcoxon_pre, 4),
    round(McNemar_pre, 4),
    paste0(round(tau_pre$estimate, 4), " (p=", round(tau_pre$p.value, 4), ")"),
    paste0(round(kappa_pre$value, 4), " (p=", round(kappa_pre$p.value, 4), ")")),
  Post = c(
    round(wilcoxon_post, 4),
    round(McNemar_post, 4),
    paste0(round(tau_post$estimate, 4), " (p=", round(tau_post$p.value, 4), ")"),
    paste0(round(kappa_post$value, 4), " (p=", round(kappa_post$p.value, 4), ")")),
  Combined = c(
    round(wilcoxon_combined, 4),
    round(McNemar_combined, 4),
    paste0(round(tau_combined$estimate, 4), " (p=", round(tau_combined$p.value, 4), ")"),
    paste0(round(kappa_combined$value, 4), " (p=", round(kappa_combined$p.value, 4), ")"))) |>
  flextable() |>
    set_caption(caption = "Malignant cases: Pre-, Post- and Combined Analysis") |>
    autofit()

```

# Table 3 and 4: Correlating data with clinical variables

```{r}
# Data preparation
data <- rawdata |> 
  # select relevant columns
  filter(complete == "yes") |>
  select(Patient_ID, Cancer, contains("Method"), -Method01_preII, Stanze_type,
    contains(c("patho_", "_stage"))) |> 
  
  # Reshape data from wide to long format
  pivot_longer(
    cols = starts_with("Method"),
    names_to = c("Method", "Timepoint"),
    names_sep = "_",
    values_to = "CTC_Count") |>
  # rename Timepoints
  mutate(Timepoint = recode(Timepoint,
                            "pre" = "CTC_pre_biopsy",
                            "post" = "CTC_post_biopsy")) |>
  pivot_wider(names_from = Timepoint,    # Reshape data back to wide format
              values_from = CTC_Count) |>
  # Compute additional columns
  mutate(CTC_total = CTC_post_biopsy + CTC_pre_biopsy,
         CTC_delta = CTC_post_biopsy - CTC_pre_biopsy,
         CTCchange = factor(case_when(
           CTC_delta > 0 ~ "CTC_inc",
           CTC_delta < 0 ~ "CTC_dec",
           CTC_delta == 0 & CTC_total != 0 ~ "stable_pos",
           TRUE ~ "stable_neg"),
           levels = c("CTC_dec", "CTC_inc", "stable_pos", "stable_neg")),
         CTCoutcome = factor(case_when(
           CTCchange %in% c("CTC_dec", "stable_neg") ~ "fav",
           CTCchange %in% c("CTC_inc", "stable_pos") ~ "unfav"),
           levels = c("fav", "unfav")),
         ER_status = factor(patho_ER, levels = c(0:12)) |>
           fct_collapse("0" = "0", other_level = ">0"),
         PR_status = factor(patho_PR, levels = c(0:12)) |>
           fct_collapse("0" = "0", other_level = ">0"),
         Ki67_status = factor(case_when(
           patho_Ki67 <= 15 ~ "low (<=15%)",
           patho_Ki67 > 15 ~ "high (>15%)"),
           levels = c("low (<=15%)", "high (>15%)")),
         'T-stage' = factor(case_when(
           T_stage <= 2 ~ "T1+T2",
           T_stage > 2 ~ "T3"),
           levels = c("T1+T2", "T3")),
         'G-stage' = factor(case_when(
           G_stage <= 2 ~ "G1+G2",
           G_stage > 2 ~ "G3"),
           levels = c("G1+G2", "G3")))
  
# data structure
flextable(tibble(Structure_data = capture.output(str(data)))) |> autofit()
# head(data, n=10) |> flextable()

```

## Result table: Correlating CTC status to clinical variables #1
```{r}

table_data <- data |> 
             filter(Method=="Method01") |>
             filter(Cancer=="negative")

variables <- c("Stanze_type","ER_status","PR_status","patho_HER2","Ki67_status",
               "T-stage","N_stage","M_stage","G-stage",
               "L_stage","V_stage","n_stage","R_stage")

table(table_data$Stanze_type,
      table_data$CTCoutcome)[c("Punch", "Vacuum"), , drop = FALSE] |>
  fisher.test()

# Loop through each variable and generate the result
result_list <- lapply(variables, function(var) {
  table(table_data[[var]], table_data$CTCoutcome) |> 
    as.data.frame.matrix() |> 
    rownames_to_column(var = var) |> 
    mutate(
      `p-value` = if_else(
        row_number() == 1,
        as.character(signif(fisher.test(table(table_data[[var]], 
                                              table_data$CTCoutcome))$p.value, 4)), ""))
  })

for (i in seq_len(min(13, length(result_list)))) {
  print(paste("Processing element", i))
  
  # Create and display flextable
  result_list[[i]] |> 
    flextable() |> 
    print()  # Print the flextable
}

```

## Result table: Correlating actual CTC counts to clinical variables #2
```{r}
# for levels 0-1 and >1
table_data <- data |>
  filter(Method == "Method02") |>
  filter(Cancer=="positive") |> 
  # for cases 0 and >0
  # mutate(CTCpool = factor(case_when(
  #           CTC_delta > 0 ~ "CTC_inc",
  #           CTC_delta < 0 ~ "CTC_dec",
  #           CTC_delta == 0 & CTC_total != 0 ~ "stable_pos",
  #            TRUE ~ "stable_neg"),
  #           levels = c("CTC_dec", "CTC_inc", "stable_pos", "stable_neg")),
  #         CTCoutcome = factor(case_when(
  #            CTCpool %in% c("stable_neg") ~ "0",
  #            CTCpool %in% c("CTC_inc", "stable_pos", "CTC_dec") ~ "!=0"),
  #            levels = c("0", "!=0")))

  # for cases 0:1 and >1
  mutate(CTCpool = factor(case_when(
            CTC_delta > 1 ~ "CTC_inc",
            CTC_delta < -1 ~ "CTC_dec",
            CTC_delta == 0 & CTC_total == 0 ~ "stable_neg",
             TRUE ~ "stable_pos"),
            levels = c("CTC_dec", "CTC_inc", "stable_pos", "stable_neg")),
          CTCoutcome = factor(case_when(
             CTCpool %in% c("CTC_dec", "CTC_inc") ~ "unfav",
             CTCpool %in% c("stable_pos", "stable_neg") ~ "fav"),
             levels = c("fav", "unfav")))

table(table_data$Stanze_type,
      table_data$CTCoutcome)[c("Punch", "Vacuum"), , drop = FALSE]

variables <- c("Stanze_type","ER_status","PR_status","patho_HER2","Ki67_status",
               "T-stage","N_stage","M_stage","G-stage",
               "L_stage","V_stage","n_stage","R_stage")

# Loop through each variable and generate the result
result_list <- lapply(variables, function(var) {
  table(table_data[[var]], table_data$CTCoutcome) |> 
    as.data.frame.matrix() |> 
    rownames_to_column(var = var) |> 
    mutate(
      `p-value` = if_else(
        row_number() == 1,
        as.character(signif(fisher.test(table(table_data[[var]], 
                                              table_data$CTCoutcome))$p.value, 4)), ""))
  })

for (i in seq_len(min(13, length(result_list)))) {
  print(paste("Processing element", i))
  
  # Create and display flextable
  result_list[[i]] |> 
    flextable() |> 
    print()  # Print the flextable
}


```

# Graphical exploration
```{r}
data |>
  filter(Method %in% c("Method01", "Method02")) |>
  ggplot(aes(x = CTC_increase, fill = Method)) +
  geom_density(alpha = 0.2) +
  labs(
    title = "Density Plot of CTC Increase by Method",
    x = "CTC Increase",
    y = "Density",
    fill = "Method"
  ) +
  theme_minimal()

# make a category with levels <= -2, -1, 0, 1, >=2 ??
# same for actual counts?
data <- 
  data |>
  mutate(CTCpre=factor(CTC_pre_biopsy,
                       levels=0:10) |> 
           fct_collapse("0"="0","1"="1",#"2"="2",#"3"="3",
                        other_level=">1"),
         CTCpost=factor(CTC_post_biopsy,
                       levels=0:10) |> 
           fct_collapse("0"="0","1"="1",#"2"="2",#"3"="3",
                        other_level=">1"))

table(data$Method, data$CTCpre)
fisher.test(data$CTCpre,data$Method)
table(data$CTCpre, data$CTCpost)
crosstab <- 
  data |> 
  filter(Method=="Method01") |> 
    group_by(CTCpre, CTCpost) |> 
  count() |> 
  pivot_wider(names_from=CTCpost, values_from=n) |> 
  mutate(across(everything(),
                ~replace_na(.x,0)))
fisher.test(crosstab[-1])
# crosstab2 <-
#   matrix(c(
#   data |> 
#   filter(Method=="Method01") |> 
#     pull(CTCpre) |>
#   table(),
#     data |> 
#   filter(Method=="Method01") |> 
#     pull(CTCpost) |>
#   table()),
#   nrow = 2,
#   byrow=T)

mcnemar.test(as.matrix(crosstab[-1]))
# Conduct paired t-test for Method A if normality is satisfied
if (shapiro_test_M01$p.value > 0.05) {
  # Paired t-test for normally distributed data
  t_test_M01 <- t.test(
    data %>% filter(Method == "Method01") %>% pull(CTC_post_biopsy),
    data %>% filter(Method == "Method01") %>% pull(CTC_pre_biopsy),
    paired = TRUE
  )
  print(t_test_M01)
} else {
  # Wilcoxon signed-rank test for non-normally distributed data
  wilcox_test_M01 <- wilcox.test(
    data %>% filter(Method == "Method01") %>% pull(CTC_post_biopsy),
    data %>% filter(Method == "Method01") %>% pull(CTC_pre_biopsy),
    paired = TRUE, correct=FALSE
  )
  print(wilcox_test_M01)
}

# paired t-test for Method B if normality is satisfied
if (shapiro_test_M02$p.value > 0.05) {
  # Paired t-test for normally distributed data
  t_test_M02 <- t.test(
    data %>% filter(Method == "Method02") %>% pull(CTC_post_biopsy),
    data %>% filter(Method == "Method02") %>% pull(CTC_pre_biopsy),
    paired = TRUE
  )
  print(t_test_M02)
} else {
  # Wilcoxon signed-rank test for non-normally distributed data
  wilcox_test_M02 <- wilcox.test(
    data %>% filter(Method == "Method02") %>% pull(CTC_post_biopsy),
    data %>% filter(Method == "Method02") %>% pull(CTC_pre_biopsy),
    paired = TRUE
  )
  print(wilcox_test_M02)
}

```

## Concordance plot

```{r}
#reliability
rawdata |> 
  filter(!is.na(Method01_pre)) |>
  ggplot(aes(Method01_pre,fill=Method01_preII))+
  geom_bar()

rawdata |> 
  filter(!is.na(Method01_pre),!is.na(Method02_pre)) |>
  ggplot(aes(Method01_pre,fill=Method02_pre))+
  geom_bar(position="fill")+
scale_y_continuous("Frequency",
                   labels=scales::percent)  

rawdata |> 
  filter(!is.na(Method01_pre),!is.na(Method02_pre)) |>
  ggplot(aes(axis1 = Method01_pre, axis2 = Method02_pre)) +
  geom_alluvium(aes(fill = Method01_pre)) +
  geom_stratum(width=.325, alpha=0) +
  labs(title = "Concordance Methods 01 and 02 pre",
       fill = "Method01_pre1") +
  #scale_fill_brewer(type = "qual", palette = 2) +
  scale_x_continuous(name = "Methods",
                   breaks=1:2,
                   labels=c("pre1","pre2"),
                   expand = expansion(mult=.01))+
  theme_minimal()
```

## Count plot pre- and post-biopsy: Style 01

```{r}
complete_cases |> 
  select(contains("Method")) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Measurement",
    values_to = "Value"
  ) |> 
  ggplot(aes(x = Measurement, y = Value)) +
  geom_beeswarm(alpha=.75) +
  labs(title = "Boxplot of Measurements for Method 01 and Method 02",
       x = "Measurement Type",
       y = "Values") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))
```

## Count plot pre- and post-biopsy: Style 02

```{r}
data_long <- complete_cases |>
  select(contains("Method"), -Method01_preII) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Measurement",
    values_to = "Value")

# Manually order the Measurement factor levels
data_long$Measurement <- factor(
  data_long$Measurement,
  levels = c("Method01_pre", "Method01_post", "Method02_pre", "Method02_post", "Method03_pre", "Method03_post"))

# Create the stacked bar chart with manual colors and custom order
ggplot(data_long, aes(x = Measurement, fill = as.factor(Value))) +
  geom_bar() +
  scale_y_continuous(
    breaks = c(0, 5, 10, 15, 20, 149, 150),
    expand = c(0, 0)) +
  scale_y_break(c(30, 149)) +
  scale_fill_brewer(palette = "Set2") +
  # scale_fill_manual(  values = c(
  #     "0" = "#D3D3D3",  # Light gray for 0 values
  #     "1" = "#1f77b4",  # Blue
  #     "2" = "#ff7f0e",  # Orange
  #     "3" = "#2ca02c",  # Green
  #     "4" = "#d62728",  # Red
  #     "5" = "#9467bd",  # Purple
  #     "6" = "#8c564b",  # Brown
  #     "7" = "#e377c2"  # Pink
  #     ),
  #   name = "Value") +
  labs(
    title = "Stacked Bar Chart ",
    x = "Measurement Type",
    y = "Count",
    fill = "Value") +
  geom_text(
    stat = "count",
    aes(label = ..count..),
    position = position_stack(vjust = 0.04),
    color = "black",
    size = 3) +
  theme_minimal() +
  theme(
    axis.line.y.right = element_blank(),
    axis.text.y.right = element_blank(),
    axis.ticks.y.right = element_blank())

```

## Count plot pre- and post-biopsy: Style 03

```{r}
df_long <- complete_cases |> 
  select(Method01_pre, Method01_post) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Measurement",
    values_to = "Value" ) |> 
  group_by(Value, Measurement) |> 
  summarise(Count = n(), .groups = 'drop')

df2_long <- expand.grid(
  Value = unique(df_long$Value),
  Measurement = c("Method01_pre", "Method01_post")) |> 
  left_join(df_long, by = c("Value", "Measurement")) |> 
  replace_na(list(Count = 0))

df2_long$Measurement <- factor(df2_long$Measurement, levels = c("Method01_pre", "Method01_post"))
ggplot(df2_long, aes(x = factor(Value), y = Count, fill = Measurement)) +
  geom_bar(stat = "identity", position = position_dodge(width = 0.9)) +
  geom_text(
    aes(label = Count),
    position = position_dodge(width = 0.8),
    vjust = -0.5,
    size = 3
  ) +
  labs(
    title = "Comparison of Method01 Pre and Post Counts",
    x = "Value",
    y = "Count",
    fill = "Measurement"
  ) +
  theme_minimal()

```

## Count plot pre- and post-biopsy: Style 04

```{r}
# Reshape the data to long format
df_long <- complete_cases |> 
  select(starts_with("Method"), -Method01_preII) |> 
  pivot_longer(
    cols = everything(),
    names_to = "Measurement",
    values_to = "Value")

# Count occurrences of each value per measurement type
df_counts <- df_long |> 
  group_by(Value, Measurement) |> 
  summarise(Count = n(), .groups = 'drop') |> 
  filter(Value != 0)

# Convert Measurement to factor variable for correct order on X-axis
df_counts$Measurement <- factor(df_counts$Measurement, levels = c("Method01_pre", "Method01_post", "Method02_pre", "Method02_post"))

# Create the stacked bar chart with counts displayed
ggplot(df_counts, aes(x = Measurement, y = Value, fill= as.factor(Count))) +
  geom_bar(stat = "identity") +
  labs(
    title = "Stacked Bar Chart of Value Counts per Measurement",
    x = "Measurement Type",
    y = "Value",
    fill = "Count") +
  scale_y_continuous(expand = c(0, 0)) +
  scale_fill_brewer(palette = "Set3") +
  theme_minimal()

```

## Plots suggested by Dr. A. Bushjan

```{r}
#reliability
complete_cases |>
  ggplot(aes(Method01_pre,fill=as.factor(Method02_pre)))+
  geom_bar()

complete_cases |>
  ggplot(aes(axis1 = Method01_pre, axis2 = Method02_pre)) +
  geom_alluvium(aes(fill = Method01_pre)) +
  geom_stratum(width=.325, alpha=0) +
  labs(title = "Concordance Method01 pre",
       fill = "Method01_pre") +
  scale_x_continuous(name = "Replicates",
                   breaks=1:2,
                   labels=c("pre1","pre2"),
                   expand = expansion(mult=.01))+
  theme_minimal()

complete_cases |>
  ggplot(aes(Method01_pre,fill=as.factor(Method02_pre)))+
  geom_bar()

complete_cases |>
  ggplot(aes(Method01_pre,fill=as.factor(Method02_pre)))+
  geom_bar(position="fill")+
scale_y_continuous("Frequency",
                   labels=scales::percent)  

complete_cases |> 
  ggplot(aes(axis1 = Method01_pre, axis2 = Method02_pre)) +
  geom_alluvium(aes(fill = as.factor(Method01_pre))) +
  geom_stratum(width=.325, alpha=0) +
  labs(title = "Concordance Methods 01 and 02 pre",
       fill = "Method01_pre") +
  scale_fill_brewer(type = "qual", palette = 2) +
  scale_x_continuous(name = "Methods",
                   breaks=1:2,
                   labels=c("pre1","pre2"),
                   expand = expansion(mult=.01))+
  theme_minimal()
```

# Plot showing the recruitment of subjects over the study period

```{r}
# cases by clinic
rawdata %>% filter(complete=="yes") |> group_by(Clinic) %>% summarise(n())
# plot
rawdata |>
  select(Date_Acrual, complete, CApositive)  |> 
  drop_na(Date_Acrual) |>
  #filter(complete.cases(Date_Acrual)) |>
  filter(complete == "yes") |>
  mutate(Individuals = seq(1:length(Date_Acrual))) |>
  mutate(Date_Acrual = as.Date(Date_Acrual)) |>
  ggplot(aes(x = sort(Date_Acrual), y = Individuals,
             color=CApositive)) + 
  geom_point(size = 2, shape = 2, alpha = 0.75) +
  # to include the smooth curve:
  # geom_smooth(method = 'lm', formula = y ~ poly(x, 2), color = 'red', se = FALSE) +
  scale_x_date(labels = date_format("%m/%Y"), date_breaks = "5 month") +
  scale_y_continuous(breaks = seq(0, 150, 20)) +
  labs(title = "Study recruitment from 05/2020 to 10/2023",
       x = "Timeline (month/Year)",
       y = "Recruited subjects",
      color = "Cancer positive") +
  theme_bw() +
  theme(
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center, bold
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold"),
    # text = element_text(size = 12),
    axis.text.x = element_text(angle = 45, hjust = 1),
    axis.ticks.x = element_line(size = 0.5)) +
  facet_grid(rows=vars(CApositive))

```

# Waterfall plot for CellSeach and Parsortix
```{r}
library(ggplot2)
library(dplyr)
library(patchwork)

# Count labeller function
count_labeller <- function(variable, value) {
  # Summarize counts for each Cancer group
  counts <- data %>%
    filter(Method == "Method02") %>%
    group_by(Cancer) %>%
    summarize(count = n(), .groups = "drop") |> 
    mutate(Cancer = case_when(
      Cancer == "positive" ~ "Cancer positive",
      Cancer == "negative" ~ "Cancer negative"
    )) |> 
    slice(c(2, 1))
  
  # Create labels with counts
  label_mapping <- setNames(paste0(counts$Cancer, " (n=", counts$count, ")"), counts$Cancer)
  
  # Return labels for the requested values
  label_mapping[value]
}

# Prepare data for plotting
plot_data <- data %>%
  filter(Method == "Method02") %>%
  mutate(CTCdelta = CTC_post_biopsy - CTC_pre_biopsy) %>%
  mutate(
    # Show actual counts where CTCdelta == 0 but CTC_total != 0
    CTCdelta = ifelse(CTCdelta == 0 & CTC_total != 0, 1, CTCdelta),
    # Set CTCdelta = 0.05 where CTC_total == 0
    CTCdelta = ifelse(CTC_total == 0, 0.05, CTCdelta)
  ) %>%
  select(Patient_ID, Cancer, Method, CTC_total, CTCdelta, CTCoutcome) %>%
  arrange(CTCdelta) %>%
  mutate(Cancer = factor(Cancer, levels = c("positive", "negative"))) %>%
  group_by(Cancer) %>%
  mutate(women = row_number())

# Create individual plots for each Cancer group
plot_positive <- plot_data %>%
  filter(Cancer == "positive") %>%
  ggplot(aes(x = women, y = CTCdelta)) +
  geom_bar(stat = "identity", width = 0.8) +
  labs(y = "CTC change from baseline", x = "Recruited women") +
  scale_x_continuous(breaks = seq(0, max(plot_data$women), by = 10)) +
  scale_y_continuous(limits = c(-3, 6)) +  # Apply consistent Y-axis limits
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    text = element_text(size = 14),
    plot.title = element_blank(),
    legend.position = "none"
  ) +
  facet_grid(cols = vars(Cancer), labeller = count_labeller)

plot_negative <- plot_data %>%
  filter(Cancer == "negative") %>%
  ggplot(aes(x = women, y = CTCdelta)) +
  geom_bar(stat = "identity", width = 0.8) +
  labs(y = NULL, x = "Recruited women") +
  scale_x_continuous(breaks = seq(0, max(plot_data$women), by = 10)) +
  scale_y_continuous(limits = c(-3, 6)) +  # Apply consistent Y-axis limits
  theme_bw() +
  theme(
    panel.grid.major.y = element_blank(),
    text = element_text(size = 14),
    plot.title = element_blank(),
    legend.position = "none"
  ) +
  facet_grid(cols = vars(Cancer), labeller = count_labeller)

# Combine plots with patchwork (60:40 ratio)
combined_plot <- plot_positive + plot_negative + 
  plot_layout(widths = c(3.3, 1.7)) +  # Set panel widths to 60:40
  plot_annotation(
    title = "Waterfall plot for Parsortix detected CTCs",
    theme = theme(plot.title = element_text(hjust = 0.5, size = 16))
  )

combined_plot

```

# Plotting CTC dynamics
```{r}
# Create the plot dataset
plot_data <- data |>
  filter(Method == "Method02") |>
  filter(Cancer=="positive") |> 
  select(Patient_ID, Cancer, Method, CTCchange, CTC_pre_biopsy, CTC_post_biopsy) |> 
  pivot_longer(cols = contains("CTC_"),
              names_to = "Timepoint",
              values_to = "CTC_count") |> 
  group_by(Patient_ID) |>
  filter(any(CTC_count != 0)) |>
  ungroup() |>
  mutate(Timepoint = factor(Timepoint, levels = c("CTC_pre_biopsy", "CTC_post_biopsy")))

# Process data and include dummy panels dynamically
panels_per_row <- 6

# Add dummy Patient_IDs if needed to balance panels
patient_ids <- unique(plot_data$Patient_ID)
num_dummies <- ceiling(length(patient_ids) / panels_per_row) * panels_per_row - length(patient_ids)

if (num_dummies > 0) {
  dummy_ids <- paste0("Dummy_", seq_len(num_dummies))
  dummy_data <- tibble(
    Patient_ID = dummy_ids,
    Cancer = NA,
    Method = "Method02",
    Timepoint = factor(c("CTC_pre_biopsy", "CTC_post_biopsy"), 
                       levels = c("CTC_pre_biopsy", "CTC_post_biopsy")),
    CTC_count = 0,
    CTCchange = NA)
  plot_data <- bind_rows(plot_data, dummy_data)
  }

# Ensure Patient_ID ordering includes dummy IDs and/or reverse row order
rows <- split(patient_ids, ceiling(seq_along(patient_ids) / panels_per_row))  # Split into rows
# ordered_ids <- unlist(rev(rows))  # Reverse rows and flatten into a single vector

plot_data <- plot_data |>
  mutate(Patient_ID = factor(Patient_ID, levels = c(patient_ids, dummy_ids)))
  # mutate(Patient_ID = factor(Patient_ID, levels = ordered_ids))

# Create the plot
ggplot(plot_data, aes(x = Timepoint, y = CTC_count, group = Patient_ID)) +
  geom_hline(yintercept = 0, linetype = "dashed", color = "black", linewidth = 0.5) + # Baseline at y = 0
  geom_line(data = plot_data |> filter(!grepl("Dummy_", Patient_ID)), alpha = 0.5, linewidth = 0.8) + # Exclude dummy panels
  
  # Pre-Biopsy Timepoint - Circles
  geom_point(
    data = plot_data |> 
      filter(Timepoint == "CTC_pre_biopsy" & !grepl("Dummy_", Patient_ID)),
    aes(x = Timepoint, y = CTC_count),
    shape = 21, size = 3, fill = "white", color = "black"
  ) +
  
  # Post-Biopsy Timepoint - Based on CTC Change
  geom_point(
    data = plot_data |> 
      filter(Timepoint == "CTC_post_biopsy" & !grepl("Dummy_", Patient_ID)),
    aes(shape = CTCchange, x = Timepoint, y = CTC_count, fill = CTCchange),
    size = 3, color = "black"
  ) +
  
  scale_shape_manual(values = c(
    "CTC_inc" = 24,   # Upward triangle
    "CTC_dec" = 25,   # Downward triangle
    "stable_pos" = 21   # Circle
  )) +
  
  scale_fill_manual(values = c(
    "CTC_inc" = "black",
    "CTC_dec" = "black",
    "stable_pos" = "white"  # Circle stays white for "no_change"
  )) +
  
  scale_x_discrete(labels = c("CTC_pre_biopsy" = "pre-biopsy", 
                              "CTC_post_biopsy" = "post-biopsy")) +
  scale_y_continuous(breaks = seq(0, 8, by = 2), limits = c(-1, 7.5)) +
  labs(
    title = "Malignant cases: Parsortix CTC dynamics compared to baseline",
    x = "Timepoint",  # Add X-axis title
    y = "CTC count"
  ) +
  theme_minimal() +
  theme(
    legend.position = "none",
    plot.title = element_text(size = 14, face = "bold", hjust = 0.5),  # Center, bold
    # X-axis title
    axis.title.x = element_text(size = 12, face = "bold", margin = margin(t = 10)),
    axis.title.y = element_text(size = 12, face = "bold"), # Y-axis title
    axis.text.x = element_text(size = 9, face = "bold", angle = 90, hjust = 1, vjust = 0.5),     # Smaller rotated X-axis labels
    axis.text.y = element_text(size = 8),   # Smaller Y-axis text
    panel.grid = element_blank(),           # Remove grid lines
    strip.text.x = element_blank(),         # Remove panel titles
    # strip.text.x = element_text(size = 10, face = "bold"),
    panel.spacing.x = unit(0.01, "lines"),   # Reduce horizontal space
    panel.spacing.y = unit(0.2, "lines"),   # Reduce vertical space
    axis.line = element_line(linewidth = 0.25),   # Add axis lines
    axis.line.x = element_blank(),          # Remove X-axis line
    axis.line.y = element_line(linewidth = 0.5),   # Retain Y-axis line
    axis.ticks.x = element_line(linewidth = 0.5), # Retain X-axis ticks
    axis.ticks = element_line(linewidth = 0.5)   # Add axis ticks
  ) +
  facet_wrap(~ Patient_ID, ncol = panels_per_row)

```


# Image Barplot
```{r}
#CS: Ca positive

df <- Ca_positive_cases %>%
  mutate(cases = c(1:97), change = BIOP2_CS - BIOP1_CS) %>%
  select(Patient_ID, BIOP1_CS, BIOP2_CS, Category_CS, cases, change)  #%>%
  filter(Category_CS == "Intact"| Category_CS == "Questionable" | Category_CS == "CEC_neg")


df_long <- pivot_longer(df, cols = starts_with("BIOP"), names_to = "time_point", values_to = "count" )
df_long<- df_long %>% mutate(count= ifelse(count==0, 0.07, count))
selected_labels <- c(1, 10, 20, 30, 40, 50, 60, 70, 80, 90, 100)

ggplot(df_long, aes(x= reorder(as.factor(cases), change),
                    y = count* ((-1)^(time_point == "BIOP2_CS")),
                    fill = as.factor(time_point))) +
  geom_bar(stat = "identity")  +
  labs(title = "Malignant (n=97)", x = "Recruited women", y = "CTC count",
       fill= "CTC detected") +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  theme(text=element_text(size=13), plot.title = element_text(hjust = 0.5), axis.ticks.x = element_blank()) +
  #scale_x_discrete(labels=as.factor(seq(1,94))) +
  scale_x_discrete(labels = ifelse(df$cases %in% selected_labels, as.character(df$cases), "")) +
  scale_fill_manual(values = c("BIOP1_CS" = "#F8766D", "BIOP2_CS" = "#00BFC4" ),
                    labels = c("pre biopsy", "post biopsy"))+
  scale_y_continuous(limits=c(-3.5, 7.5), breaks=seq(-4,7,by=1))


#CS: Ca negative
df <- Ca_negative_cases %>%
  mutate(cases = c(1:50), change = BIOP2_CS - BIOP1_CS) %>%
  select(Patient_ID, BIOP1_CS, BIOP2_CS, Category_CS, cases, change)  #%>%
  #filter(Category_CS == "Intact"| Category_CS == "Questionable" | Category_CS == "CEC_neg")


df_long <- pivot_longer(df, cols = starts_with("BIOP"), names_to = "time_point", values_to = "count" )
df_long<- df_long %>% mutate(count= ifelse(count==0, 0.07, count))
selected_labels <- c(1, 10, 20, 30, 40, 50)
ggplot(df_long, aes(x= reorder(as.factor(cases), change),
                    y = count* ((-1)^(time_point == "BIOP1_CS")),
                    fill = as.factor(time_point))) +
  geom_bar(stat = "identity")  +
  labs(title = "Benign (n=50)", x = "Recruited women", y = "CTC count",
       fill= "CTC detected") +
  theme_bw() +
  theme(panel.grid.minor.y = element_blank()) +
  theme(text=element_text(size=11), plot.title = element_text(hjust = 0.5), axis.ticks.x = element_blank()) +
  #scale_x_discrete(labels=as.factor(seq(1,93))) +
  scale_x_discrete(labels = ifelse(df$cases %in% selected_labels, as.character(df$cases), "")) +
  scale_fill_manual(values = c("BIOP2_CS" = "#00BFC4", "BIOP1_CS" = "#F8766D" ),
                    labels = c("post biopsy", "pre biopsy"))+
  scale_y_continuous(limits=c(-3.5, 7.5), breaks=seq(-3.5,7,by=1))


```

